---
title: "Assignment 1"
author: "Catherine Saint"
date: "8/13/2020"
output: html_document
---

```{r, message = FALSE}
library(tidycensus)
library(tidyverse)
```

## Variables used: 

I decided to explore employment trends at the census tract level in Los Angeles County, California (so my unit of analysis is the census tract). I chose the following variables from ACS-5 year summary 2018. The reason behind this choice is because, according to the ACS/API handbook, estimates for small geographic areas with populations less than 20,000 people are only available for the ACS5 (five-year data).
https://www.census.gov/content/dam/Census/library/publications/2020/acs/acs_api_handbook_2020.pdf


1. Median Age (continuous) : med_age

2. Part of the total population unemployed in civilian labor force (age 16 years and above) (continuous): unemp_labor

3. Part of the total population who are high school graduates (or equivalent) in labor force (continuous): 

4. Majority race of the population (not Hispanic or Latino) (categorical) : maj_race

5. Whether the majority of the population is foreign born or not(categorical): maj_foreign

```{r}
pop_vars <- c(med_age =           "B01002_001",
              unemp_labor =       "B23025_005",
              hs_grad =           "B23006_010", 
              total_pop =         "B02001_001", 
              white_alone =       "B03002_003",
              black_AA_alone =    "B03002_004",
              native_alone =      "B03002_005",
              asian_alone =       "B03002_006",
              hawaiian_PI_alone = "B03002_007",
              other =             "B03002_008",
              mult_races =        "B03002_009",
              hispanic_latino =   "B03002_012",
              foreign_born =      "B05002_013")

tractLosAngelesCA2018 <- get_acs(geography= "tract", 
              variables = pop_vars, 
              output = "wide", 
              state = "CA",
              county = "Los Angeles",
              year = 2018, 
              survey = "acs5")
```

## Calculating new variables:

After finding my variables and filtering them to census tracts in Los Angeles county, I want to calculate the percentage of people unemployed in each census tract, and the percentage of people with a highschool degree or equivalent. I also calculate the percentage of the population that is foreign born, and create a categorical variable with two categories: "TRUE" if at least  half the population are foreign born, and "FALSE" otherwise. 

I leave median income as is, as there is nothing to change. I also pull up detailed variables for my race categorical variable and manipulate it so that any tract with a race greater than 50% will be designated as the majority race.  

```{r}
tractLosAngelesCA2018 <- tractLosAngelesCA2018 %>%
  mutate(pct_white = white_aloneE/total_popE,
         pct_black = black_AA_aloneE/total_popE,
         pct_hisp_lat = hispanic_latinoE/total_popE,
         pct_native = native_aloneE/total_popE,
         pct_asian = asian_aloneE/total_popE,
         pct_hawaiian_PI = hawaiian_PI_aloneE/total_popE,
         pct_mult = mult_racesE/total_popE,
         pct_other = otherE/total_popE,
         pct_unemp_labor = unemp_laborE / total_popE,
         pct_hs_grad = hs_gradE / total_popE, 
         maj_foreign = foreign_bornE / total_popE > 0.5 )%>%
 mutate(maj_race = case_when(pct_white > 0.5 ~ "nh_white",
                             pct_black > 0.5 ~ "nh_black",
                             pct_hisp_lat > 0.5 ~ "hs_latino",
                             pct_native > 0.5 ~ "native_am",
                             pct_asian > 0.5 ~ "nh_asian",
                             pct_hawaiian_PI > 0.5 ~ "hawaiian_PI",
                             pct_mult > 0.5 ~ "multi",
                             pct_other > 0.5 ~ "other_nh",
                             TRUE ~ "no_majority")) %>%
  select (maj_foreign, maj_race, pct_unemp_labor, 
          pct_hs_grad, med_ageE)
  
```

Now that I have calculated my variables, I'll display the first few rows of the data frame. 

```{r}
tractLosAngelesCA2018
```

My final dataset has 2,346 rows, which mean my analysis will include 2,346 census tracts.

Finally, I'll save my data as a CSV file so I can use it later.

```{r}
write_csv(tractLosAngelesCA2018,"tractLosAngelesCA2018.csv")
```
