---
title: "Assignment 3"
author: "Catherine Saint"
date: "9/15/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggplot2)
tract_data <- read_csv("tractLosAngelesCA2018.csv")
```

### Topic Reminder

I want to explore the relationship between age, unemployment, educational attainment, race, and nativity on the census tract level in Los Angeles County. 

## Relationships between age, unemployment, and educational attainment

First, I will test the relationship between my continuous variables - median age, percent unemployed, and percent highschool graduates - by running a correlation test. 

```{r cars}
correlation1 <- cor.test(tract_data$med_ageE, tract_data$pct_unemp_labor)
correlation1

correlation2 <- cor.test(tract_data$med_ageE, tract_data$pct_hs_grad)
correlation2

correlation3 <- cor.test(tract_data$pct_unemp_labor, tract_data$pct_hs_grad)
correlation3
```

The following statements can be made by looking at Pearson's r and p-values for the correlation tests above. 

  - median age and unemployment: a weak, but statistically signficant, negative correlation at a 95% confidence level. 
  
  - median age and high school graduate: a weak, but statistically significant, negative correlation at a 95% confidence level.
  
  - unemployment and highschool Ggraduate: a weak, but statistically significant, positive correlation at a 95% confidence level. 
  
  
I want to  see what these "weaker" relationships look like, so I chose median Age and unemployment to represent on a scatterplot. 


```{r}
ggplot(tract_data, aes(x = pct_unemp_labor, y = med_ageE)) + 
    geom_point(size = 0.5, alpha = 0.25, position = "jitter") +
    geom_smooth() +
    theme_bw() +
    scale_x_continuous(name = "Unemployment (Percent)") +
    scale_y_continuous(name = "Median Age",
                      breaks = seq(0, 105, by = 5), 
                      labels = paste("",
                                  prettyNum(seq(0, 105, by = 5),
                                            big.mark = ","))) 
```

## Relationship between race and nativity

I use the chi-square test of association to determine whether there is a statistically signficant association between race and nativity. 
```{r}
chi_sq <- chisq.test(tract_data$maj_race, tract_data$maj_foreign == "TRUE")
chi_sq
chi_sq$observed
chi_sq$expected
```

I see that the Native American category violates the chi-square observed table, as it has a value less than 5. However, with a p-value so much less than 0.05, we can assume a statistically significant relationship between race and nativity at a 95% confidence level despite this possible innacuracy. 

## Relationships between race and continuous variables 

I use the anova test to find the relationship between race and my continuous variables: age, unemployment, and educational attainment. 

```{r}
anova <- aov (med_ageE ~ maj_race, data = tract_data)
summary(anova)

anova <- aov (pct_hs_grad ~ maj_race, data = tract_data)
summary(anova)

anova <- aov (pct_unemp_labor ~ maj_race, data = tract_data)
summary(anova)

```

Difference Check:
```{r}
differences <- TukeyHSD(anova)
as_tibble(cbind(pair = row.names(differences$maj_race),
                differences$maj_race))
```


There exists a significant association between race and unemployment, race and median age, and race and education attainment level at a 95% confidence level. All p-values fall well below the 0.05 threshold. 

I chose to display the relationship between median age and majority race using a box-plot. 
```{r}
ggplot(tract_data, aes(x = maj_race, y = med_ageE)) +
  geom_boxplot() +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90)) +
  scale_x_discrete(name = "Majority Race") +
  scale_y_continuous(name = "Median Age",
                     breaks = seq(0, 105, by = 5),
                     labels = paste(prettyNum(seq(0, 105, by = 5), 
                                              big.mark = ",")))
```


## Relationships between nativity and continuous variables 

I use a two-sample t-test to test the relationship between nativity and my continuous variables: age, unemployment, and educational attainment.

```{r}
difference1 = t.test(med_ageE ~ maj_foreign == "TRUE", 
                    data = tract_data)
difference1
difference2 = t.test(pct_unemp_labor ~ maj_foreign == "TRUE", 
                    data = tract_data)
difference2
difference3 = t.test(pct_hs_grad ~ maj_foreign == "TRUE", 
                    data = tract_data)
difference3

```
There is no signficant relationship found between nativity and median age at a 95% confidence level. 

There is a significant relationship found between unemployment and nativity at a 95% confidence level; tracts with a majority foreign born population have a higher unemployment percentage. 

There is a strong, significant relationship found between educational attainment and nativity; tracts with a majority foreign born population have a larger percentage of highschool graduate (or equivalent) educational attainment than do majority native born tracts. 
