---
title: "Assignment 1"
author: "Cat"
date: "8/13/2020"
output: html_document
---

```{r}
library(tidycensus)
library(tidyverse)
census_api_key
```

## Finding Variables 

### Area-level data (American Community Survey)
```{r}
area_vars_2018 <- load_variables(2018, "acs5")
```
## Variables used: 

I decided to explore employment trends by census tract on the county level (unit of analysis) in Los Angeles, California. I chose the following variables from ACS-5 year summary 2018. The reason behind this choice is because, according to the ACS/API handbook, estimates for small geographic areas with populations less than 20,000 people are only available for the ACS5 (five-year data).
https://www.census.gov/content/dam/Census/library/publications/2020/acs/acs_api_handbook_2020.pdf


1. Median Age (continuous) : med_age

2. Part of the total population unemployed in civilian labor force (age 16 years and above) (continuous): unemp_labor

3. Part of the total population who are high school graduates (or equivalent) in labor force (continuous): 

4. Majority race of the population (not Hispanic or Latino) (categorical) : maj_race

5. Whether the majority of the population is foreign born or not(categorical): maj_foreign

```{r}
pop_vars <- c(med_age = "B01002_001",
              unemp_labor = "B23025_005",
              hs_grad = "B23006_010", 
              total_pop = "B02001_001", 
              white_alone = "B03002_003",
              black_AA_alone = "B03002_004",
              hispanic_latino = "B03002_012",
              native_alone = "B03002_005",
              asian_alone = "B03002_006",
              hawaiian_PI_alone = "B03002_007",
              mult_races = "B03002_009",
              other = "B03002_008",
              foreign_born = "B05002_013")

tractLosAngelesCA2018 <- get_acs(geography= "tract", 
              variables = pop_vars, 
              output = "wide", 
              state = "CA",
              county = "Los Angeles",
              year = 2018, 
              survey = "acs5")
```
After finding my variables and filtering them to census tracts in Los Angeles county, I want to calculate the percentage of people unemployed in each census tract, and the percentage of people with a highschool degree or equivalent. I also calculate the percentage of the majority foreign born, and set it so that if half the population are foreign born, then the census tract is labeled as true. 

I leave median income as is, as there is nothing to change. I also pull up detailed variables for my race categorical variable and manipulate it so that any tract with a race greater than 50% will be designated as the majority race.  
```{r}
tractLosAngelesCA2018 <- tractLosAngelesCA2018 %>%
  mutate(pct_white = white_aloneE/total_popE,
         pct_black = black_AA_aloneE/total_popE,
         pct_hisp_lat = hispanic_latinoE/total_popE,
         pct_native = native_aloneE/total_popE,
         pct_asian = asian_aloneE/total_popE,
         pct_hawaiian_PI = hawaiian_PI_aloneE/total_popE,
         pct_mult = mult_racesE/total_popE,
         pct_other = otherE/total_popE,
         pct_unemp_labor = unemp_laborE / total_popE,
         pct_hs_grad = hs_gradE / total_popE, 
         maj_foreign = foreign_bornE / total_popE > 0.5 )%>%
  
  mutate(pct_nh_white = pct_white,
         pct_nh_black = pct_black,
         pct_nh_asian = pct_asian,
         pct_other = pct_native + pct_hawaiian_PI + pct_mult + pct_other) %>%
  
  mutate(maj_race_tract = case_when(
    pct_nh_white > 0.5 ~ "nh_white",
    pct_nh_black > 0.5 ~ "nh_black",
    pct_hisp_lat > 0.5 ~ "hs_latino",
    pct_nh_asian > 0.5 ~ "nh_asian",
    pct_other > 0.5 ~ "other",
    TRUE ~ "no_majority")) %>%
  
  select (maj_foreign, maj_race_tract, pct_unemp_labor, pct_hs_grad, med_ageE)
  
```
Now that I have calculated my variables, I want to select them and create a data frame, which I will then convert to a CSV. 
```{r}
tractLosAngelesCA2018
write_csv(tractLosAngelesCA2018,"tractLosAngelesCA2018.csv")

```